### 2020-05-31 writeup/proof_examples.tex
sec 3.3

@@

%--------old arguments
%With the same arguments as in Lemma \ref{lem_hat_v} we can show that \eqref{hatv_add1} holds, and under the choice of parameters \eqref{choiceofpara},
%\be \nonumber
%\begin{split}
%g(M)&=(1+\OO(n^{-\e}))\cdot d^2 \tr\left[(X_1^{\top}X_1)(X_1^{\top}X_1 +X_2^{\top}X_2)^{-1}\Sigma_2(X_1^{\top}X_1 +X_2^{\top}X_2)^{-2} (X_1^{\top}X_1)\right] \\ 
%&+(1+\OO(n^{-\e}))\cdot \sigma^2  \bigtr{\Sigma_2(X_1^{\top}X_1  + X_2^{\top}X_2)^{-1} }  =: \delta_\beta(M)+ g_{\vari}(M),
%\end{split}
%\ee
%with high probability for any constant $\e>0$. In fact, Lemma \ref{lem_hat_v} was proved assuming that $M=\id$, but its proof can be easily extended to the case with general $M\in \cal S_{\mu}$ by using that $\lambda(M)\in [\mu_{\min},\mu_{\max}]$. We omit the details here. 
%
%
%With Lemma \ref{lem_cov_shift}, we can obtain that with high probability,
%\begin{align}\label{gvar_extra}
%g_{\vari}(M_0)= \frac{\sigma^2}{\rho_1+\rho_2}\cdot \frac1p\tr\left( \frac{1}{a_1M_0^\top M_0 + a_2}\right)  +\OO(p^{-1/2+\e}),
%\end{align}
%where $(a_1,a_2)$ satisfies \eqref{eq_a12extra}. On the other hand, by Theorem \ref{thm_model_shift} we can estimate $\delta_\beta$ as
%\be \label{deltabeta_extra}
%\begin{split}
%&\left| \delta_{\beta}(M_0) - \frac{d^2\cdot \rho_1^2}{(\rho_1 + \rho_2)^2} \tr\left[\Sigma_1^{1/2}{M_0} \frac{(1 + a_3)\id + a_4 {M_0}^{\top}{M_0}}{(a_2 + a_1 {M_0}^{\top}{M_0})^2} {M_0}^{\top}\Sigma_1^{1/2}\right]\right| \le \delta + \OO(p^{1-\e}d^2),
%\end{split}
%\ee
%where $(a_3,a_4)$ satisfies \eqref{eq_a34extra} and $\delta$ is defined in \eqref{eq_deltaextra}.
%
%
%Next we derive some estimates for $a_i$, $i=1,2,3,4$. From equation \eqref{eq_a12extra}, it is easy to see that
%\be\label{est_a12extra}a_1\ge \frac{\rho_1-1}{\rho_1+\rho_2},\quad a_2\le \frac{\rho_2}{\rho_1+\rho_2}.\ee
%Then solving \eqref{eq_a34extra}, we get that
%\begin{align}
%0\le a_3&= \frac{ b_0 (\rho_1a_1^{-2}-b_2) + b_1^2 }{  \left(\rho_2  a_2^{-2}-  b_0\right)\left(\rho_1 a_1^{-2} -  b_2  \right) - b_1^2} \le \frac{ b_0 \rho_1a_1^{-2}  }{  \rho_1\rho_2  a_1^{-2}a_2^{-2}-  b_0\rho_1 a_1^{-2} - b_2\rho_2  a_2^{-2}}\nonumber\\
%&\le \frac{\rho_1}{\rho_1\rho_2-\rho_1 -\rho_2 },\label{est_a3extra}
%\end{align}
%and
%\begin{align}
%0\le a_4&= \frac{b_1\rho_2 a_2^{-2}}{\left(\rho_2  a_2^{-2}-  b_0\right)\left(\rho_1 a_1^{-2} -  b_2  \right) - b_1^2} \le  \frac{ b_1\rho_2 a_2^{-2} }{  \rho_1\rho_2  a_1^{-2}a_2^{-2}-  b_0\rho_1 a_1^{-2} - b_2\rho_2  a_2^{-2}} \nonumber\\
%&\le  \frac{ \rho_2 (b_1a_1^{2}) }{  \rho_1\rho_2 -  \rho_1  - \rho_2 } \le \frac{\rho_2\cdot \mu_{\min}^{-2}}{\rho_1\rho_2-\rho_1-\rho_2}=\OO\left( \frac{1}{\rho_1}\right),\label{est_a4extra}
%\end{align}
%where in the above derivations we also used that $b_1^2\le b_0 b_2$ by Cauchy-Schwarz inequality, $b_0 a_2^2 < 1$, $b_2 a_1^2 < 1$, and $b_1 a_1^2 \le \mu_{\min}^{-2}$.
%Inserting \eqref{est_a12extra} into \eqref{gvar_extra}, we obtain that with high probability,
%\begin{align}\label{approximateteM}
%\le g(M_0) \le  \frac{\sigma^2}{\rho_1+\rho_2}\cdot \frac1p\tr\left( \frac{1}{a_1M_0^\top M_0}\right) \cdot \left(1 +\OO(p^{-\e})\right)
%\end{align}
%where
%$$\wt g(M_0):=\frac{\sigma^2}{\rho_1+\rho_2}\cdot \frac1p\tr\left( \frac{1}{a_1M_0^\top M_0}\right)  +\frac{pd^2\cdot \rho_1^2}{(\rho_1 + \rho_2)^2} \cdot \frac1p\tr\left(\Sigma_1\frac{1 + a_3}{a_1^2 M_0^{\top}{M_0}} \right),$$
%and the error satisfies
%$$|\cal E|\le C\left( \frac{\rho_2}{\rho_1} +\rho_1^{-1/2} +\OO(p^{-\e})\right)\wt g(M_0).$$
%Here the constant $C>0$ depends only on $\mu_{\max}$, $\mu_{\min}$ and $\|\Sigma_1\|$, but otherwise does not depend on $\rho_1$ and $\rho_2$. 
%--------end old argument---------

>>>>

To compare different choices of $M$ on the performance of $\hat{\beta}_t^{\MTL}$, consider the following family of matrices
\begin{align*}
		\cS_{\mu}\define\bigset{M \mid \det\bigbrace{ M^\top M} \le \mu^p, \lambda(M) \in [\mu_{\min}, \mu_{\max}]},
\end{align*}
	where $\mu, \mu_{\min}, \mu_{\max}$ are fixed values that do not grow with $p$.
	We assume that $\beta_1$ and $\beta_2$ are generated following the isotropic model with $d = 0$.
The following proposition shows that when $n_1$ is much larger than $n_2$, $\te(\hat{\beta}^{\MTL})$ is minimized at $M = \mu\id$ among all matrices in $\cS_{\mu}$.

\begin{proposition}[Covariate shift]\label{prop_covariate}
	In the setting described above, assume that $\rho_1, \rho_2>1$.
	Let $g(M)$ denote the test error of $\hat{\beta}_t^{\MTL}$ when the covariance shift matrix is equal to $M\in\cS_{\mu}$.
	We have that \[ g(\mu\id) \le \bigbrace{1+ \bigo{\frac{\rho_2}{\rho_1}  }} \min_{M\in\cS_{\mu}} g(M). \]
\end{proposition}
Proposition \ref{prop_covariate} implies that when $\rho_1\gg \rho_2$, having no covariate shift is the optimal choice for choosing the source task.
%This provides evidence that covariate shift is unfavorable when there are many source task datapoints,

%\todo{} To complement the result, we show an example when the statement is not true if $n_1 \le n_2$.

@@

%--------old arguments
%With the same arguments as in Lemma \ref{lem_hat_v} we can show that \eqref{hatv_add1} holds, and under the choice of parameters \eqref{choiceofpara},
%\be \nonumber
%\begin{split}
%g(M)&=(1+\OO(n^{-\e}))\cdot d^2 \tr\left[(X_1^{\top}X_1)(X_1^{\top}X_1 +X_2^{\top}X_2)^{-1}\Sigma_2(X_1^{\top}X_1 +X_2^{\top}X_2)^{-2} (X_1^{\top}X_1)\right] \\ 
%&+(1+\OO(n^{-\e}))\cdot \sigma^2  \bigtr{\Sigma_2(X_1^{\top}X_1  + X_2^{\top}X_2)^{-1} }  =: \delta_\beta(M)+ g_{\vari}(M),
%\end{split}
%\ee
%with high probability for any constant $\e>0$. In fact, Lemma \ref{lem_hat_v} was proved assuming that $M=\id$, but its proof can be easily extended to the case with general $M\in \cal S_{\mu}$ by using that $\lambda(M)\in [\mu_{\min},\mu_{\max}]$. We omit the details here. 
%
%
%With Lemma \ref{lem_cov_shift}, we can obtain that with high probability,
%\begin{align}\label{gvar_extra}
%g_{\vari}(M_0)= \frac{\sigma^2}{\rho_1+\rho_2}\cdot \frac1p\tr\left( \frac{1}{a_1M_0^\top M_0 + a_2}\right)  +\OO(p^{-1/2+\e}),
%\end{align}
%where $(a_1,a_2)$ satisfies \eqref{eq_a12extra}. On the other hand, by Theorem \ref{thm_model_shift} we can estimate $\delta_\beta$ as
%\be \label{deltabeta_extra}
%\begin{split}
%&\left| \delta_{\beta}(M_0) - \frac{d^2\cdot \rho_1^2}{(\rho_1 + \rho_2)^2} \tr\left[\Sigma_1^{1/2}{M_0} \frac{(1 + a_3)\id + a_4 {M_0}^{\top}{M_0}}{(a_2 + a_1 {M_0}^{\top}{M_0})^2} {M_0}^{\top}\Sigma_1^{1/2}\right]\right| \le \delta + \OO(p^{1-\e}d^2),
%\end{split}
%\ee
%where $(a_3,a_4)$ satisfies \eqref{eq_a34extra} and $\delta$ is defined in \eqref{eq_deltaextra}.
%
%
%Next we derive some estimates for $a_i$, $i=1,2,3,4$. From equation \eqref{eq_a12extra}, it is easy to see that
%\be\label{est_a12extra}a_1\ge \frac{\rho_1-1}{\rho_1+\rho_2},\quad a_2\le \frac{\rho_2}{\rho_1+\rho_2}.\ee
%Then solving \eqref{eq_a34extra}, we get that
%\begin{align}
%0\le a_3&= \frac{ b_0 (\rho_1a_1^{-2}-b_2) + b_1^2 }{  \left(\rho_2  a_2^{-2}-  b_0\right)\left(\rho_1 a_1^{-2} -  b_2  \right) - b_1^2} \le \frac{ b_0 \rho_1a_1^{-2}  }{  \rho_1\rho_2  a_1^{-2}a_2^{-2}-  b_0\rho_1 a_1^{-2} - b_2\rho_2  a_2^{-2}}\nonumber\\
%&\le \frac{\rho_1}{\rho_1\rho_2-\rho_1 -\rho_2 },\label{est_a3extra}
%\end{align}
%and
%\begin{align}
%0\le a_4&= \frac{b_1\rho_2 a_2^{-2}}{\left(\rho_2  a_2^{-2}-  b_0\right)\left(\rho_1 a_1^{-2} -  b_2  \right) - b_1^2} \le  \frac{ b_1\rho_2 a_2^{-2} }{  \rho_1\rho_2  a_1^{-2}a_2^{-2}-  b_0\rho_1 a_1^{-2} - b_2\rho_2  a_2^{-2}} \nonumber\\
%&\le  \frac{ \rho_2 (b_1a_1^{2}) }{  \rho_1\rho_2 -  \rho_1  - \rho_2 } \le \frac{\rho_2\cdot \mu_{\min}^{-2}}{\rho_1\rho_2-\rho_1-\rho_2}=\OO\left( \frac{1}{\rho_1}\right),\label{est_a4extra}
%\end{align}
%where in the above derivations we also used that $b_1^2\le b_0 b_2$ by Cauchy-Schwarz inequality, $b_0 a_2^2 < 1$, $b_2 a_1^2 < 1$, and $b_1 a_1^2 \le \mu_{\min}^{-2}$.
%Inserting \eqref{est_a12extra} into \eqref{gvar_extra}, we obtain that with high probability,
%\begin{align}\label{approximateteM}
%\le g(M_0) \le  \frac{\sigma^2}{\rho_1+\rho_2}\cdot \frac1p\tr\left( \frac{1}{a_1M_0^\top M_0}\right) \cdot \left(1 +\OO(p^{-\e})\right)
%\end{align}
%where
%$$\wt g(M_0):=\frac{\sigma^2}{\rho_1+\rho_2}\cdot \frac1p\tr\left( \frac{1}{a_1M_0^\top M_0}\right)  +\frac{pd^2\cdot \rho_1^2}{(\rho_1 + \rho_2)^2} \cdot \frac1p\tr\left(\Sigma_1\frac{1 + a_3}{a_1^2 M_0^{\top}{M_0}} \right),$$
%and the error satisfies
%$$|\cal E|\le C\left( \frac{\rho_2}{\rho_1} +\rho_1^{-1/2} +\OO(p^{-\e})\right)\wt g(M_0).$$
%Here the constant $C>0$ depends only on $\mu_{\max}$, $\mu_{\min}$ and $\|\Sigma_1\|$, but otherwise does not depend on $\rho_1$ and $\rho_2$. 
%--------end old argument---------
we observe the following two phases as we increase $n_1 / p$ in Figure \ref{fig_covariate}.
When $n_1 \le n_2$, having complementary covariance matrices leads to lower test error compared to the case when $\Sigma_1 = \Sigma_2$.
When $n_1 > n_2$, having complementary covariance matrices leads to higher test error compared to the case when $\Sigma_1 = \Sigma_2$.
We provide a theoretical justification below.

### 2020-05-31 writeup/transfer.tex
sec 3.3

\begin{algorithm}[!t]
	\caption{An incremental training schedule for multi-task learning over two tasks}
	\label{alg_inc_train}
	\begin{algorithmic}[1]
		\Input Two tasks $(X_1, Y_1)$ and $(X_2, Y_2)$.
		\Param A shared module $B$, output layers $W_1, W_2$ as in the hard parameter sharing architecture.
		\Req Batch size $S$, number of epochs $T$, validation accuracy $\hat{g}(B; W_1, W_2)$.
		\State Divide $(X_1, Y_1)$ randomly into $K$ batches of size $S$ as $(x^{(1)}, y^{(1)}), \dots, (x^{(K)}, y^{(K)})$.
		\For{$i = 1,\dots, K$}
			\For{$j = 1,\dots, T$}
				\State Update $B, W_1, W_2$ using the training data $\set{x^{(j)}, x^{(j)}}_{j=1}^i$ and  $(X_2, Y_2)$.
			\EndFor
		\State Let $a_i = \hat{g}(B; W_1, W_2)$ be the validation loss after epoch $i$.
			If $a_i < a_{i+1}$, \textbf{break}.
		\EndFor
	\end{algorithmic}
\end{algorithm}

@@

So far we have considered the isotropic model where $\Sigma_1 = \Sigma_2$.
This setting is relevant for settings where different tasks share the same input features such as multi-class image classification.
In general, the covariance matrices of the two tasks may be different such as in text classification.
In this part, we consider what happens when $\Sigma_1 \neq \Sigma_2$.

We measure covariate shift by the matrix $M = \Sigma_1^{1/2} \Sigma_2^{-1/2}$.
We ask: is it better to have $M$ as being close to identity, or should $M$ involve varying levels of singular values?
Understanding this question has implications for applying normalization methods in multi-task learning \cite{LV19,CBLR18,YKGLHF20}.
We show that if $n_1$ is much larger than $n_2$, then the optimal $M$ matrix should be proportional to identity, under certain assumptions on its range of singular values (to be formulated in Proposition \ref{prop_covariate}).
On the other hand, if $n_1$ is comparable or even smaller than $n_2$, we show an example where having ``complementary'' covariance matrices is better performing than having the same covariance matrices.

	To compare different choices of $M$ on the performance of $\hat{\beta}_t^{\MTL}$, consider the following family of matrices
	\begin{align*}
		\cS_{\mu}\define\bigset{M \mid \det\bigbrace{ M^\top M} \le \mu^p, \lambda(M) \in [\mu_{\min}, \mu_{\max}]},
	\end{align*}
	where $\mu, \mu_{\min}, \mu_{\max}$ are fixed values that do not grow with $p$.
	We assume that $\beta_1$ and $\beta_2$ are generated following the isotropic model with $d = 0$.
	The following proposition shows that when $n_1$ is much larger than $n_2$, $\te(\hat{\beta}^{\MTL})$ is minimized at $M = \mu\id$ among all matrices in $\cS_{\mu}$.

\begin{proposition}[Covariate shift]\label{prop_covariate}
	In the setting described above, assume that $\rho_1, \rho_2>1$.
	Let $g(M)$ denote the test error of $\hat{\beta}_t^{\MTL}$ when the covariance shift matrix is equal to $M\in\cS_{\mu}$.
	We have that \[ g(\mu\id) \le \bigbrace{1+ \bigo{\frac{\rho_2}{\rho_1}  }} \min_{M\in\cS_{\mu}} g(M). \]
\end{proposition}
Proposition \ref{prop_covariate} implies that when $\rho_1\gg \rho_2$, having no covariate shift is the optimal choice for choosing the source task.
%This provides evidence that covariate shift is unfavorable when there are many source task datapoints,

\todo{} To complement the result, we show an example when the statement is not true if $n_1 \le n_2$.
\begin{example}\label{ex_complement}
	In the setting of Proposition \ref{prop_covariate}, we compare two cases: (i) when $M = \id$; (ii) when $M$ has $p/2$ singular values that are equal to $\lambda$ and $p/2$ singular values that are equal to $1 / \lambda$.
	For simplicity, we assume that $d = 0$.
	Hence the two tasks have the same model parameters.

	In Figure \ref{fig_model_shift_phasetrans} (c), we plot the test error of the target task for $n_2 = 4p$ and $n_1$ ranging from $p$ to $20p$.
	Second, we observe the following two phases as we increase $n_1 / p$.
	When $n_1 \le n_2$, having complementary covariance matrices leads to lower test error compared to the case when $\Sigma_1 = \Sigma_2$.
	When $n_1 > n_2$, having complementary covariance matrices leads to higher test error compared to the case when $\Sigma_1 = \Sigma_2$.
	A theoretical justification can be found in Appendix \ref{app_proof_31}.
\end{example}

>>>>

\begin{algorithm}[!t]
	\caption{An incremental training schedule for multi-task learning over two tasks}
	\label{alg_inc_train}
	\begin{algorithmic}[1]
		\Input Two tasks $(X_1, Y_1)$ and $(X_2, Y_2)$.
		\Param A shared module $B$, output layers $W_1, W_2$ as in the hard parameter sharing architecture.
		\Req Batch size $S$, number of epochs $T$, validation accuracy $\hat{g}(B; W_1, W_2)$.
		\State Divide $(X_1, Y_1)$ randomly into $K$ batches of size $S$ as $(x^{(1)}, y^{(1)}), \dots, (x^{(K)}, y^{(K)})$.
		\For{$i = 1,\dots, K$}
			\For{$j = 1,\dots, T$}
				\State Update $B, W_1, W_2$ using the training data $\set{x^{(j)}, x^{(j)}}_{j=1}^i$ and  $(X_2, Y_2)$.
			\EndFor
		\State Let $a_i = \hat{g}(B; W_1, W_2)$ be the validation accuracy after epoch $i$.
			If $a_i < a_{i-1}$, \textbf{break}.
		\EndFor
	\end{algorithmic}
\end{algorithm}

@@

So far we have considered the isotropic model where $\Sigma_1 = \Sigma_2$.
This setting is relevant for settings where different tasks share the same input features such as multi-class image classification.
In general, the covariance matrices of the two tasks may be different such as in text classification.
In this part, we consider what happens when $\Sigma_1 \neq \Sigma_2$.
We show that when $n_1 / n_2$ is large, MTL with covariate shift can be suboptimal compared to MTL without covariate shift.

\begin{example}\label{ex_complement}
	We measure covariate shift by the matrix $M = \Sigma_1^{1/2} \Sigma_2^{-1/2}$.
	Assume that $\Psi(\beta_1, \beta_2) = 0$ for simplicity.
	We compare two cases: (i) when $M = \id$; (ii) when $M$ has $p/2$ singular values that are equal to $\lambda$ and $p/2$ singular values that are equal to $1 / \lambda$.
	Hence, $\lambda$ measures the severity of covariate shift.
	Figure \ref{fig_covariate} shows a simulation of this setting by varying $\lambda$.
\end{example}
%We ask: is it better to have $M$ as being close to identity, or should $M$ involve varying levels of singular values?
%Understanding this question has implications for applying normalization methods in multi-task learning \cite{LV19,CBLR18,YKGLHF20}.
%We show that if $n_1$ is much larger than $n_2$, then the optimal $M$ matrix should be proportional to identity, under certain assumptions on its range of singular values (to be formulated in Proposition \ref{prop_covariate}).
%On the other hand, if $n_1$ is comparable or even smaller than $n_2$, we show an example where having ``complementary'' covariance matrices is better performing than having the same covariance matrices.
